---
title: "Weight Lifting Exercise"
author: "Jose Vicente Carrasquero Aumaitre"
date: "September 27, 2015"
output: 
  html_document: 
    theme: journal
---

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

We begin installing needed packages and loading libraries. We include Random Forest for classification and regression, Regressive Partitioning and Regression Trees and Decision Tree plot.
We set seed to guarantee reproduceability

```{r}
library(caret)
library(randomForest) 
library(rpart)
library(rpart.plot)  
set.seed(1234)
```

Now we are loading both training and testing data and setting missing values to NA. Then we delete columns with all missing values and delete variables irrelevant to the project.

```{r, echo=FALSE}
trainingset <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!", ""))
testingset <- read.csv('pml-testing.csv', na.strings=c("NA","#DIV/0!", ""))
trainingset<-trainingset[,colSums(is.na(trainingset)) == 0]
testingset <-testingset[,colSums(is.na(testingset)) == 0]
trainingset <-trainingset[,-c(1:7)]
testingset <-testingset[,-c(1:7)]
```

Let's create subtraining and subtesting datasets 

```{r}
subsamples <- createDataPartition(y=trainingset$classe, p=0.60, list=FALSE)
subTraining <- trainingset[subsamples, ] 
subTesting <- trainingset[-subsamples, ]
```

Bar Plot of levels of the variable classe within the subTraining data set and trining the Model 1
```{r}
plot(subTraining$classe, col="blue", main="Bar Plot of levels of the variable classe within the subTraining data set", xlab="classe levels", ylab="Frequency")

model1 <- rpart(classe ~ ., data=subTraining, method="class")
```

Applying the prediction process with model 1 and plotting of the Decision Tree

```{r}
prediction1 <- predict(model1, subTesting, type = "class")

rpart.plot(model1, main="Classification Tree", extra=102, under=TRUE, faclen=0)
```

Running the Confusion Matrix for Prediction 1
```{r}
confusionMatrix(prediction1, subTesting$classe)
```

Creating and testing Model 2 by using Random Forest and subtesting data set

```{r}
model2 <- randomForest(classe ~. , data=subTraining, method="class")

prediction2 <- predict(model2, subTesting, type = "class")

confusionMatrix(prediction2, subTesting$classe)
```

